MS_CNuptake.Rmd

### 2022-07-11 MS_CNuptake.Rmd
chapter 1 MS code

# Setup 
```{r setup, include=FALSE}
rm(list=ls())
library(MLmetrics)
library(tidyverse) 
library(ncmeta)
library(viridis)
library(ggthemes)
library(LSD)
library(yardstick)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
library(gplots)
library(tidyselect)
library(extrafont)
library(caret)
library(recipes)
#devtools::load_all("/Users/yunpeng/yunkepeng/rbeni/")
library(rbeni)
library(raster)
library(maps)
library(rworldmap)
library(cowplot)
library(ncdf4)
library(scales)
library(lme4)
library(lmerTest)
library("PerformanceAnalytics")
library(MuMIn)
library(tidyverse)
library(ggplot2)
library(lme4)
library(visreg)
library(ggpubr)
library(car)
library("ggplotify")
library(remotes)
library(tune)
library(relaimpo)

@@

### BP Again ??? XXX
```{r}
#now, start works
BP_dataset2 <- na.omit(NPP_forest[,c("tnpp_a","age_a","fAPAR_a","CNrt_a","Tg_a","PPFD_a","vpd_a","site_a")])
dim(BP_dataset2)
a2 <- stepwise(BP_dataset2,"tnpp_a")
a2[[1]]
a2[[2]]
a2[[3]]
bp_model <- (lmer(tnpp_a~Tg_a+fAPAR_a+PPFD_a+CNrt_a+age_a+(1|site_a),data=BP_dataset2))
summary(bp_model)
r.squaredGLMM(bp_model)

@@

### ANPP again? XXX
```{r}
anpp_tnpp_dataset2 <- na.omit(NPP_forest[,c("anpp_tnpp_a","age_a","fAPAR_a","CNrt_a","Tg_a","PPFD_a","vpd_a","site_a")])
dim(subset(NPP_forest,ANPP_2>0))
model2a <- stepwise(anpp_tnpp_dataset2,"anpp_tnpp_a")
model2a[[1]]
model2a[[2]]
model2a[[3]]
anpp_tnpp_model <- (lmer(anpp_tnpp_a~CNrt_a+PPFD_a+Tg_a+age_a+(1|site_a),data=anpp_tnpp_dataset2))
summary(anpp_tnpp_model)
r.squaredGLMM(anpp_tnpp_model)

@@

### Root C:N
XXX why is variable called 'forest'?
```{r warning=FALSE}
# Root C/N = 94 - as derived from median values of collected samples
NPP_forest_sitemean <- aggregate(NPP_forest,by=list(NPP_forest$site), FUN=mean, na.rm=TRUE) 
summary(NPP_forest_sitemean$CN_root_final) # using median = 94
dim(subset(NPP_forest,CN_root_final>0))

@@

## NRE

XXX this is about NRE, but section 'NRE' above as well 
```{r}
#NRE = 69%
NRE_Du <- read.csv(file="~/data/NRE_various/NRE_Du/NRE_Du.csv")
NRE_Dong <- read.csv(file="~/data/NRE_various/NRE_Deng/NRE_Deng.csv")

@@

# XXX explain how hard-coded numbers are obtained. Better replace them by dynamically defined variables. 
NPP_forest$pred_wnf <- NPP_forest$pred_wnpp/100
NPP_forest$pred_bnf <- NPP_forest$pred_bnpp/94

@@

source("R/calc_area.R")
lonlat <- vcmax25_df[,c("lon","lat")]
area_m2 <- calc_area(lonlat$lat,0.5,0.5)
#fland - to show each grid's land cover percentage
nc <- read_nc_onefile("~/data/fland/global.fland.nc") #Input nc
output_fland <- nc_to_df(nc, varnam = "fland")
fland <- output_fland$myvar
#include conversion factor (from g to Pg)
conversion <- area_m2 * fland /1e+15
```

@@

### ???
XXX what is this figure?
```{r}
gg <- plot_map3(na.omit(nue_all[,c("lon","lat","nue_CNrt")]),varnam = "nue_CNrt",latmin = -65, latmax = 85,combine=FALSE,
                colorscale = c( "royalblue4", "wheat","tomato3"),
                breaks = seq(-0.05,0.05,0.01))
g1 <- gg$ggmap +labs(title = "Soil C/N")+theme_grey(base_size = 15);g2 <- gg$gglegend

@@

### ???
XXX what is this figure?
```{r warning=FALSE}
#figs2 representing all predictors
gg <- plot_map3(na.omit(CNrt[,c("lon","lat","CNrt")]),varnam = "CNrt",latmin = -65, latmax = 85,combine=FALSE)
d1 <- gg$ggmap +labs(title = "Soil C/N")+theme_grey(base_size = 15)+ white;d2 <- gg$gglegend+ white

@@

## Read data
```{r}

#trendy
#model output
#CABLE-POP
CABLE_GPP <- as.data.frame(nc_to_df(read_nc_onefile("~/data/trendy/v8/CABLE-POP_S2_gpp_ANN_mean.nc"), varnam = "gpp"))
CABLE_NPP <- as.data.frame(nc_to_df(read_nc_onefile("~/data/trendy/v8/CABLE-POP_S2_npp_ANN_mean.nc"), varnam = "npp"))

@@

ISAM_fNup <- as.data.frame(nc_to_df(read_nc_onefile("~/data/trendy/v8/ISAM_S2_fNup_ANN_mean.nc"), varnam = "fNup"))

# XXX what are these numbers that you multiply here?
ISAM_fNup$myvar<- ISAM_fNup$fNup*1000*31556952

ISAM_gpp <- as.data.frame(nc_to_df(read_nc_onefile("~/data/trendy/v8/ISAM_S2_gpp_ANN_mean.nc"), varnam = "gpp"))
ISAM_gpp$myvar<- ISAM_gpp$gpp*1000*31556952

ISAM_npp <- as.data.frame(nc_to_df(read_nc_onefile("~/data/trendy/v8/ISAM_S2_npp_ANN_mean.nc"), varnam = "npp"))
ISAM_npp$myvar<- ISAM_npp$npp*1000*31556952

@@

## Plot???

XXX What plot?
```{r}

@@

allmaps<- list(CABLE_GPP,CABLE_NPP,ISAM_fNup,ISAM_gpp,ISAM_npp,ISBA_GPP,ISBA_NPP,
               JULES_GPP,JULES_NPP,LPJ_GPP,LPJ_NPP,ORCHIDEE_GPP,ORCHIDEE_NPP,
               ORCHICNP_fNup,ORCHICNP_GPP,ORCHICNP_NPP,SDGVM_GPP,SDGVM_NPP)
obj_name <- c("CABLE_GPP","CABLE_NPP","ISAM_fNup","ISAM_gpp","ISAM_npp","ISBA_GPP","ISBA_NPP",
              "JULES_GPP","JULES_NPP","LPJ_GPP","LPJ_NPP","ORCHIDEE_GPP","ORCHIDEE_NPP",
              "ORCHICNP_fNup","ORCHICNP_GPP","ORCHICNP_NPP","SDGVM_GPP","SDGVM_NPP")
#calculate nue 
summary( ISAM_npp$myvar/ISAM_fNup$myvar)
summary(na.omit(ORCHICNP_NPP$myvar/ORCHICNP_fNup$myvar))

@@

## XXX ???
```{r}

@@

## XXX ??
```{r}

@@

## XXX ?
```{r}
# n uptake
mod_nue3 <- (lm(pred_nuptake~LMA_a+fAPAR_a+age_a+vpd_a+CNrt_a+Tg_a+vcmax25_a+PPFD_a,data=nue_allplots))
summary(mod_nue3)
nue1c <- visreg(mod_nue3,"LMA_a",type="contrast")
nue2c <- visreg(mod_nue3,"fAPAR_a",type="contrast")
nue3c <- visreg(mod_nue3,"age_a",type="contrast")
nue4c <- visreg(mod_nue3,"vpd_a",type="contrast")
nue5c <- visreg(mod_nue3,"CNrt_a",type="contrast")
nue6c <- visreg(mod_nue3,"Tg_a",type="contrast")
nue7c <- visreg(mod_nue3,"vcmax25_a",type="contrast")
nue8c <- visreg(mod_nue3,"PPFD_a",type="contrast")

@@

## XXX ???
```{r}
#validation - N uptake
Nmin_statistical$pred_nuptake

@@

## XXX what figure does this produce? 
```{r}
#visreg

@@

### XXX?
```{r}
#all statistics
#forest
sum(uncertainty_npp*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_anpp*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_bnpp*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_lnpp*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_wnpp*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_lnf*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_wnf*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_bnf*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_nuptake*(forest_percent *conversion)*available_grid2,na.rm=TRUE)

@@

### XXX?
```{r}
#partial residual figure
BP_dataset <- na.omit(NPP_forest[,c("tnpp_a","obs_age_a","observedfAPAR_a","soilCN_a","Tg_a","PPFD_a","vpd_a","site_a")])
#model1 <- stepwise(BP_dataset,"tnpp_a")
#model1[[1]]
#model1[[2]]
bp_model <- (lmer(tnpp_a~Tg_a+PPFD_a+soilCN_a+obs_age_a+observedfAPAR_a+(1|site_a),data=BP_dataset))
summary(bp_model)

>>>>

# Setup 
```{r setup, include=FALSE}
rm(list=ls())
library(MLmetrics)
library(vhs)
library(tidyverse) 
library(ncmeta)
library(viridis)
library(ggthemes)
library(LSD)
library(yardstick)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
library(gplots)
library(tidyselect)
library(extrafont)
library(caret)
library(recipes)
#remotes::install_github("cj-holmes/vhs")
devtools::load_all("/Users/yunpeng/yunkepeng/rbeni_new/rbeni/")
library(rbeni)
library(raster)
library(maps)
library(rworldmap)
library(cowplot)
library(ncdf4)
library(scales)
library(lme4)
library(lmerTest)
library("PerformanceAnalytics")
library(MuMIn)
library(tidyverse)
library(ggplot2)
library(lme4)
library(visreg)
library(ggpubr)
library(car)
library("ggplotify")
library(remotes)
library(tune)
library(relaimpo)

@@

### BP vs. model predictors
```{r}
#now, start works
BP_dataset2 <- na.omit(NPP_forest[,c("tnpp_a","age_a","fAPAR_a","CNrt_a","Tg_a","PPFD_a","vpd_a","site_a")])
dim(BP_dataset2)
a2 <- stepwise(BP_dataset2,"tnpp_a")
a2[[1]]
a2[[2]]
a2[[3]]
bp_model <- (lmer(tnpp_a~Tg_a+fAPAR_a+PPFD_a+CNrt_a+age_a+(1|site_a),data=BP_dataset2))
summary(bp_model)
r.squaredGLMM(bp_model)

@@

### ANPP vs. modelled predictors
```{r}
anpp_tnpp_dataset2 <- na.omit(NPP_forest[,c("anpp_tnpp_a","age_a","fAPAR_a","CNrt_a","Tg_a","PPFD_a","vpd_a","site_a")])
dim(subset(NPP_forest,ANPP_2>0))
model2a <- stepwise(anpp_tnpp_dataset2,"anpp_tnpp_a")
model2a[[1]]
model2a[[2]]
model2a[[3]]
anpp_tnpp_model <- (lmer(anpp_tnpp_a~CNrt_a+PPFD_a+Tg_a+age_a+(1|site_a),data=anpp_tnpp_dataset2))
summary(anpp_tnpp_model)
r.squaredGLMM(anpp_tnpp_model)

@@

### Root C:N
```{r warning=FALSE}
# Root C/N = 94 - as derived from median values of collected samples
NPP_forest_sitemean <- aggregate(NPP_forest,by=list(NPP_forest$site), FUN=mean, na.rm=TRUE) 
summary(NPP_forest_sitemean$CN_root_final) # using median = 94
dim(subset(NPP_forest,CN_root_final>0))

@@

## NRE 
```{r}
#NRE = 69%
NRE_Du <- read.csv(file="~/data/NRE_various/NRE_Du/NRE_Du.csv")
NRE_Dong <- read.csv(file="~/data/NRE_various/NRE_Deng/NRE_Deng.csv")

@@

# 100 and 94 are constant.

NPP_forest$pred_wnf <- NPP_forest$pred_wnpp/100
NPP_forest$pred_bnf <- NPP_forest$pred_bnpp/94

@@

source("R/calc_area.R")
lonlat <- vcmax25_df[,c("lon","lat")]
area_m2 <- calc_area(lonlat$lat,0.5,0.5)
#fland - to show each grid's land cover percentage
nc <- read_nc_onefile("~/data/fland/global.fland.nc") #Input nc
output_fland <- nc_to_df(nc, varnam = "fland")
fland <- output_fland$fland
#include conversion factor (from g to Pg)
conversion <- area_m2 * fland /1e+15
```

@@

### global maps for effect of each factor NUE (not used now)
```{r}
gg <- plot_map3(na.omit(nue_all[,c("lon","lat","nue_CNrt")]),varnam = "nue_CNrt",latmin = -65, latmax = 85,combine=FALSE,
                colorscale = c( "royalblue4", "wheat","tomato3"),
                breaks = seq(-0.05,0.05,0.01))
g1 <- gg$ggmap +labs(title = "Soil C/N")+theme_grey(base_size = 15);g2 <- gg$gglegend

@@

### Global maps of each predictors
```{r warning=FALSE}
#figs2 representing all predictors
gg <- plot_map3(na.omit(CNrt[,c("lon","lat","CNrt")]),varnam = "CNrt",latmin = -65, latmax = 85,combine=FALSE)
d1 <- gg$ggmap +labs(title = "Soil C/N")+theme_grey(base_size = 15)+ white;d2 <- gg$gglegend+ white

@@

## Read data
```{r}
#trendy
#model output
#CABLE-POP
CABLE_GPP <- as.data.frame(nc_to_df(read_nc_onefile("~/data/trendy/v8/CABLE-POP_S2_gpp_ANN_mean.nc"), varnam = "gpp"))
CABLE_NPP <- as.data.frame(nc_to_df(read_nc_onefile("~/data/trendy/v8/CABLE-POP_S2_npp_ANN_mean.nc"), varnam = "npp"))

@@

ISAM_fNup <- as.data.frame(nc_to_df(read_nc_onefile("~/data/trendy/v8/ISAM_S2_fNup_ANN_mean.nc"), varnam = "fNup"))
ISAM_fNup$fNup<- ISAM_fNup$fNup*1000*31556952

ISAM_gpp <- as.data.frame(nc_to_df(read_nc_onefile("~/data/trendy/v8/ISAM_S2_gpp_ANN_mean.nc"), varnam = "gpp"))
ISAM_gpp$gpp<- ISAM_gpp$gpp*1000*31556952

ISAM_npp <- as.data.frame(nc_to_df(read_nc_onefile("~/data/trendy/v8/ISAM_S2_npp_ANN_mean.nc"), varnam = "npp"))
ISAM_npp$npp<- ISAM_npp$npp*1000*31556952

@@

## TRENDY analysis- BP
```{r}

@@

allmaps<- list(CABLE_GPP,CABLE_NPP,ISAM_fNup,ISAM_gpp,ISAM_npp,ISBA_GPP,ISBA_NPP,
               JULES_GPP,JULES_NPP,LPJ_GPP,LPJ_NPP,ORCHIDEE_GPP,ORCHIDEE_NPP,
               ORCHICNP_fNup,ORCHICNP_GPP,ORCHICNP_NPP,SDGVM_GPP,SDGVM_NPP)
obj_name <- c("CABLE_GPP","CABLE_NPP","ISAM_fNup","ISAM_gpp","ISAM_npp","ISBA_GPP","ISBA_NPP",
              "JULES_GPP","JULES_NPP","LPJ_GPP","LPJ_NPP","ORCHIDEE_GPP","ORCHIDEE_NPP",
              "ORCHICNP_fNup","ORCHICNP_GPP","ORCHICNP_NPP","SDGVM_GPP","SDGVM_NPP")
#calculate nue 
summary( ISAM_npp$npp/ISAM_fNup$ISAM_npp)
summary(na.omit(ORCHICNP_NPP$npp/ORCHICNP_fNup$ISAM_npp))

@@

## TRENDY analysis - N uptake
```{r}

@@

## TRENDY analysis - NNUE
```{r}

@@

## Nuptake or NUE vs. all predictors
```{r}
# n uptake
mod_nue3 <- (lm(pred_nuptake~LMA_a+fAPAR_a+age_a+vpd_a+CNrt_a+Tg_a+vcmax25_a+PPFD_a,data=nue_allplots))
summary(mod_nue3)
nue1c <- visreg(mod_nue3,"LMA_a",type="contrast")
nue2c <- visreg(mod_nue3,"fAPAR_a",type="contrast")
nue3c <- visreg(mod_nue3,"age_a",type="contrast")
nue4c <- visreg(mod_nue3,"vpd_a",type="contrast")
nue5c <- visreg(mod_nue3,"CNrt_a",type="contrast")
nue6c <- visreg(mod_nue3,"Tg_a",type="contrast")
nue7c <- visreg(mod_nue3,"vcmax25_a",type="contrast")
nue8c <- visreg(mod_nue3,"PPFD_a",type="contrast")

@@

## N uptake validation between models
```{r}
#validation - N uptake
Nmin_statistical$pred_nuptake

@@

## visreg for BP and anpp/bp vs. all measured predictors
```{r}
#visreg

@@

### statistics of uncertainty
```{r}
#all statistics
#forest
sum(uncertainty_npp*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_anpp*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_bnpp*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_lnpp*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_wnpp*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_lnf*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_wnf*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_bnf*(forest_percent *conversion)*available_grid2,na.rm=TRUE)
sum(uncertainty_nuptake*(forest_percent *conversion)*available_grid2,na.rm=TRUE)

@@

### BP partial relationship vs. observations
```{r}
#partial residual figure
BP_dataset <- na.omit(NPP_forest[,c("tnpp_a","obs_age_a","observedfAPAR_a","soilCN_a","Tg_a","PPFD_a","vpd_a","site_a")])
#model1 <- stepwise(BP_dataset,"tnpp_a")
#model1[[1]]
#model1[[2]]
bp_model <- (lmer(tnpp_a~Tg_a+PPFD_a+soilCN_a+obs_age_a+observedfAPAR_a+(1|site_a),data=BP_dataset))
summary(bp_model)
